---
title: "capsule measurement"
output: html_notebook
---

## Goal

To measure the amount (intensity and/or thickness) of capsule in each individual cell and relate this to the fluorescence intensity in this given cell. I was not extremely happy with the outcome of a few segmentation programs so I decided to go oldschool for segmentation by phase and then measurement of the surroundings by flourescence in R. I start this with a "test dataset" of only one strain, if it works, I will redo the analysis on all 6 strains I have (always on/off controls, etc).

## Test dataset

* Morphometrics output of phase contrast segmentation of image stack VL4315
* Image stack VL4315, phase contrast
* Image stack VL4315, capsule (antibody)
* Image stack VL4315, mScarlet
* Image stack VL4315, YFP
* Image stack VL4315, mTurquoise

## Preparation

```{r}

library(bactMAP)
library(dplyr)
library(ggplot2)

setwd("C:/Users/adminuser/Documents/Microscopy Files/Anne-Stephanie/capsulequantification/20210901_capsulequantification/VL4315")

addPixels2um("0.05", 0.05) #the lif file I have has this registered as the pixel conversion factor

#get files (don't need phase)
morph <- extr_Morphometrics("phase_02-Sep-2021_CONTOURS.mat", mag="0.05")
capsule <- extr_OriginalStack("capsule.tif")
mScarlet <- extr_OriginalStack("mSc.tif")
YFP <- extr_OriginalStack("YFP.tif")
mTurquoise <- extr_OriginalStack("mTQ.tif")
```

## Analysis

I decided that the easiest way to go is to take a certain amount of pixels/micron outside the phase-contrast segmentation. The nice thing is also that in this way, you won't have to do another segmentation step and just can use the analysis in one go. 

The idea:

* Take & convert phase-contrast data
* Use the function gBuffer (rgeos package) to add a ring of x micron or pixels to each cell.
* Use extr_OriginalCells to get the pixel values inside this larger ring. 
* Remove all pixels that are inside the smaller ring (the cell).
* Compute 1) the average intensity of the pixels (+sd, etc) 2) 

```{r}

## function for making a region around the cell


makeBuffer <- function(dataset, width_pixels){
  dataset$unique <- paste(dataset$frame, dataset$cell, sep="_")
  out <- lapply(unique(dataset$unique), function(W){
    rgeos::gBuffer(sp::SpatialPointsDataFrame(coords=dataset[dataset$unique==W,][,c("X", "Y")], data=dataset[dataset$unique==W,]), width= width_pixels)@polygons[[1]]@Polygons[[1]]@coords |>
      data.frame() |> 
      mutate(cell = unique(dataset$cell[dataset$unique==W]),
             frame = unique(dataset$frame[dataset$unique==W]))
  }) |> 
    bind_rows() |>
    rename(X=x, Y=y) |>
    left_join(unique(dataset[,c("cell", "frame", "area", "pole1", "pole2", "max.length", "max.width", "angle", "Xmid", "Ymid", "maxwum", "max_um", "area_um")]))
  return(out)
}

#dataset = morph$mesh
#width_pixels = 0.1 / morph$pixel2um
buffered <- makeBuffer(morph$mesh, width_pixels = 0.2/morph$pixel2um)

#take values of each cell
cellsbuffered <- extr_OriginalCells(capsule, buffered, turnCell=FALSE) |> mutate(values = values - min(values))
cells_mSc <- extr_OriginalCells(mScarlet, morph$mesh, turnCell=FALSE) |> mutate(values = values - min(values))
cells_YFP <- extr_OriginalCells(YFP, morph$mesh, turnCell=FALSE) |> mutate(values = values - min(values))
cells_mTQ <- extr_OriginalCells(mTurquoise, morph$mesh, turnCell=FALSE) |> mutate(values = values - min(values))

#remove inside of buffered cell
onlybuffer <- cellsbuffered |> anti_join(select(cells_mSc, -values)) #note that this could have been YFP, mTQ as well. it's just the mask.
onlybuffer_onePC <- onlybuffer |>
  select(-x, -y) |>
  onePerCell() |> 
  rename(mean_capsule = mean,
         sd_capsule = sd)

bgbuf <- quantile(cellsbuffered$values, 0.7)

onlybuffercount <- onlybuffer[onlybuffer$values>bgbuf,] |> group_by(cell) |>  summarize(capsulesize = n()) |> full_join(onlybuffer_onePC)
allcolors <- combineDataframes(list(cells_mSc, cells_YFP, cells_mTQ), listofchannels=list("mScarlet", "YFP", "mTurquoise"))$finalframe |> 
  group_by(cell, frame, channel) |> summarize(values_mean = mean(values),
                                              values_sd = sd(values),
                                              cellsizepix = n()) |>
  left_join(onlybuffercount) |> mutate(capsize_cor = capsulesize/cellsizepix)

```


```{r}
plotcors <- ggplot(allcolors, aes(x=capsize_cor, y=values_mean, color=channel)) + 
  geom_point() + geom_smooth(method="lm", fill=NA) + 
  scale_x_log10() + 
  scale_y_log10() + 
  ggpubr::stat_regline_equation(aes(label=..rr.label..)) + 
  scale_color_manual(values=c("mScarlet" = "#e71e28", "YFP" = "#c69448", "mTurquoise" = "#52a5bd")) + 
  theme_minimal() + 
  xlab("capsule size (pixels)") + 
  ylab("mean fluorescence intensity inside cell (AU)")

ggsave(plotcors, width=10, height=10, filename="FPvsCapsule.pdf")
save(allcolors, file = "allcolors.Rda")
save(morph, file="phase_morphometrics.Rda")
save(cells_mSc, file= "cells_mScarlet.Rda")
save(cells_YFP, file = "cells_YFP.Rda")
save(cells_mTQ, file= "cells_mTurquoise.Rda")
save(onlybuffer, file= "cells_buffered.Rda")
```

